\LARGE \bf Mean Estimation from Adaptive One-bit Measurements

### 2017-10-05 Slides/OneBit_Allerton_slides_V2.tex
new version of presentation

\begin{frame}
\frametitle{Efficiency}
\textbf{Definition:} \emph{asymptotic relative efficiency (ARE)}:
\[
\mathrm{ARE}(\widehat{\theta}_n) \triangleq \lim_{n\rightarrow \infty} \mathbb E \left[ \left(\widehat{\theta}_n-\theta \right)^2 \right] / \left(\sigma^2 / n \right)
\]
(relative to minimax risk without bit constraint)

%SAY: many definitions for ARE are possible, we use this one becasue our results apply to all of them

@@

\begin{frame}
\frametitle{ARE under Centralized Encoding}

\begin{prop} ARE under centralized encoding is $1$
\end{prop}
\pause

\textbf{Proof:}
\[
\mathbb E \left( \theta- \widehat{\theta} \right)^2 = \overbrace{\mathbb E \left( \theta- \bar{\theta} \right)^2}^{\sigma^2/n} + \underbrace{\mathbb E \left( \bar{\theta}- \widehat{\theta} \right)^2}_{O(2^{-2n})}
\]
\pause
\begin{itemize}
\item Encoder is required to describe $\bar{\theta}$ using $n$ bits
\begin{itemize}
\item divide parameter space $\Theta$ into $2^n$ regions of equal size
\item send region index where $\bar{\theta}$ falls
\end{itemize}
\end{itemize}
\bigskip
\pause
Note: mean of $\bar{\theta}$ is unknown -- hard to derive a globally optimal strategy
%SAY: exploration -- exploitation even in this simple setting
\end{frame}

@@

\begin{center}
\vspace{-20pt}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]
  \node at (0,0) (source) {$\thetac_j$} ;
  \node[ssum, right of = source, scale = 0.75] (sum1) {+};
  \node[above of = sum1, node distance = 1cm] (z1) {$Z_{1}$};
  \node[int1, right of = sum1, node distance = 2cm, scale = 0.75] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (sum1); 
\draw[->,line width = 2pt] (sum1) -- node[above] {$X_{1}$} (enc1); 
\draw[->,line width = 2pt] (z1) -- (sum1);

@@

%\node[int1, right of = enc2, node distance = 6cm, scale = 0.75] (est) {Est};
\draw[->,line width = 2pt] (enc1) -- node[above, xshift = 0cm, rotate = -15] (mes1) {one bit} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm, rotate = 12] (mes2) {one bit} (est);   
\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\thetac}_j$};
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\pause
Assume:
\begin{itemize}
\item $Z_{1},\ldots,Z_{n} \overset{\mathrm{i.i.d}}{\sim}\mathcal N(0,\sigma^2)$
\pause
\item Replicate $k$ times: $\thetac_1,\ldots,\thetac_k \overset{\iid}{\sim} \mathcal N(0,\sigma_0^2)$ 
\pause
\item Block encode $X_{j,1},\ldots,X_{j,n}$ using $k$ bits, $j=1,\ldots,k$
\pause
\item $D_{CEO} = \inf_{k,\widehat{\theta}} \frac{1}{k} \sum_{j=1}^k \mathbb E \left( \theta_j - \widehat{\theta}_j\right)^2$
\end{itemize}
%SAY: lower bound to our setting
\end{frame}

\begin{frame}
Quadratic Gaussian CEO under optimal rate allocation [Chen, Zhang, Berger, Wicker '04] :
\[
 D_{CEO}  \geq \frac{4}{3} \frac{\sigma^2}{n} + o(1/n) 
\]
%\begin{itemize}
%\item[] CEO distortion satisfies [Chen, Zhang, Berger, Wicker '04]:
%\[ D^\star = \frac{4}{3} %\frac{\sigma^2}{n} + o(n^{-1}) \]
%\end{itemize}
\pause
\begin{alertblock}{Conclusion}
Distributed encoding will hurt you (even if you can repeat experiment and encode over blocks)
\end{alertblock}
\bigskip
\pause
\begin{itemize}
\item In fact [K., Rini, Goldsmith '17]:
\end{itemize} 
\[
 D_{CEO} \leq \frac{4}{3} \frac{\sigma^2}{n} + \frac{\sigma_0^2}{3n} + o(1/n) 
\]
(ARE of $4/3$ is achievable with coding over blocks)
%SAY: depending on how large our bound from $k=1$, it tells us that distributed setting is not the limitng factor
\end{frame}

\section{Adaptive Encoding}

@@

\begin{itemize}
\item Distributed: $M_i = f_i(X_i)$  \hfill{ARE $\geq 4/3$}
\visible<2->{
\item {Centralized}: $M^n = f(X_1,\ldots,X_n)$ \hfill{ARE $=1$} }
\visible<3->{
\item {\color{red} Adaptive / Sequential: $M_i = f_i(X_i,M^{i-1})$ \hfill{ARE$=\pi/2$} }  }
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item[(i)] For $X \sim \mathcal{N}(\theta,\sigma^2)$, $\mathsf{med}(X) = \theta$ 
\pause
\item[(ii)] $\mathsf{med}(X) = \underset{m}{ \mathrm{argmin~}} \mathbb E \left|X-m \right|$
\pause
\item[(iii)] Stochastic gradient descent on $\mathbb E \left| X-\theta \right|$:
\[
\theta_n = \theta_{n-1} + \gamma_n \sgn(X_n - \theta_n)
\]
\[
\widehat{\theta}_n = \frac{1}{n} \sum_{i=1}^n \theta_i 
\]
\end{itemize}
\bigskip
\pause
From [Polyak \& Juditsky '92] (under conditions on $(\gamma_n)$): 
\[
\sqrt{n}(\theta- \widehat{\theta}_n) \rightarrow \mathcal N\left(0, \sigma^2 \pi /2 \right)
\]
\end{frame}
%picture

@@

\begin{frame}
\frametitle{Converse}
\framesubtitle{ARE $\geq \pi/2$}
The van-Trees inequality (e.g. [Tsybakov '08]) implies
\[
\mathbb E \left(\theta - \widehat{\theta} \right)^2 \geq \frac{1}{\mathbb E I_\theta(M^n) + I_\pi} \geq \frac{1}{\sum_{i=1}^n I_\theta(M_i|M^{i-1}) + I_\pi} 
\]
$I_\pi$ is the location Fisher information with w.r.t. some prior $\pi(d\theta)$ on $\Theta$\\

@@

\begin{frame}
\frametitle{Distributed Encoding}
\framesubtitle{Threshold Detection}
We consider only messages of the form
\[
M_i = \sgn(X_i - t_i),\quad i=1,\ldots,n
\]
\begin{center}
\pause
\begin{tikzpicture}
\node[coordinate] at(0,0) (ti) {};
\draw[color = red, line width = 1.5pt] (ti)+(0,0.2) -- node[above, yshift = 0.07cm] {\color{black} $t_i$} +(0,-0.2);
\draw[<-,line width = 1pt] (-3,0) -- node[above] {$M_i=-1$} (ti);
\draw[->,line width = 1pt] (ti) -- node[above] {$M_i=1$} (6,0);
\end{tikzpicture}
\end{center}
Assume:
\[
\lambda_n([a,b]) = \frac{1}{n} \mathsf{card}\left( [a,b]  \cap \left\{t_i \right\} \right)
\]
converges weakly to a probability distribution $\lambda$
\\
\bigskip
\pause
\textbf{Example:} $t_1,\ldots,t_n$ drawn i.i.d. from a distribution $\lambda$ on $\mathbb R$

@@

\begin{frame}
\frametitle{Interpretations}
\begin{itemize}
\pause
\item ML estimator is local asymptotically minimax 
\pause
\item ARE of ML is $1/K_{\lambda}(\theta)$ -- 
only depends on asymptotic threshold density $\lambda$
\pause
\item  
\[
\mathrm{ARE} = \frac{1}{K_{\lambda}(\theta)}  = \frac{\sigma^2}{\int \eta \left( \frac{t-\theta}{\sigma}\right) \lambda(dt)}
> \pi / 2
\]
(equality iff $\lambda = \delta_{\theta}$)
%SAY: cannot be as good as sequential thresholds
%\item Threshold density $\lambda$ that maximizes
\pause
\item Minimax $\lambda$ can be found using a convex program -- minimax ARE  depends on radius of $\Theta$
%ARE depends on size of parameter space $\Theta$ 
%\begin{center}
%\begin{tikzpicture}
%\node at (0,0)  {\includegraphics[scale=0.3]{k_vs_b2}};
%\node at (0,-1.7) {$r$};
%\node[rotate = 90] at (-2.3,0) {$1/K_{\lambda^\star}$};
%\end{tikzpicture}
%
%Minimax ARE vs radius $r$ of parameter space
%\end{center}

@@

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\pause
\item Centralized encoding: ARE$=1$
\pause
\item Adaptive:
\begin{itemize}
\item ARE is $\pi/2$ (regardless of radius of $\Theta$)
\pause
\item $\sim 1.57$ more samples are required due to 1-bit constraints
\end{itemize}
\pause
%\pause
%\item One-step optimal one-bit message is a threshold detector
\bigskip
\item Distributed threshold detection:
\begin{itemize}
\item ARE of MLE characterized by density of threshold values 
\item MLE is local asymptotically optimal 
\item minimax ARE depends on radius of $\Theta$
\end{itemize}
%\item Minimax ARE of MLE depends on size of parameter space
\end{itemize}

@@

>>>>

\begin{frame}
\frametitle{Efficiency}
\textbf{Definition:} \emph{asymptotic relative efficiency (ARE)}:
\[
\mathrm{ARE}(\widehat{\theta}_n) \triangleq \lim_{n\rightarrow \infty} \mathbb E \left[ \left(\widehat{\theta}_n-\theta \right)^2 \right] / \left(\sigma^2 / n \right)
\]
(relative to minimax risk without bit constraint)
%SAY: many definitions for ARE are possible, we use this one becasue our results apply to all of them
\bigskip

\pause
\begin{prop} ARE under centralized encoding is $1$
\end{prop}
\pause
\textbf{Proof:}
\[
\mathbb E \left( \theta- \widehat{\theta} \right)^2 = \overbrace{\mathbb E \left( \theta- \bar{\theta} \right)^2}^{\sigma^2/n} + \underbrace{\mathbb E \left( \bar{\theta}- \widehat{\theta} \right)^2}_{O(2^{-2n})}
\]

@@

%\begin{frame}
%\frametitle{ARE under Centralized Encoding}
%
%\begin{prop} ARE under centralized encoding is $1$
%\end{prop}
%\pause
%
%\textbf{Proof:}
%\[
%\mathbb E \left( \theta- \widehat{\theta} \right)^2 = \overbrace{\mathbb E \left( \theta- \bar{\theta} \right)^2}^{\sigma^2/n} + \underbrace{\mathbb E \left( \bar{\theta}- \widehat{\theta} \right)^2}_{O(2^{-2n})}
%\]
%\pause
%\begin{itemize}
%\item Encoder is required to describe $\bar{\theta}$ using $n$ bits
%\begin{itemize}
%\item divide parameter space $\Theta$ into $2^n$ regions of equal size
%\item send region index where $\bar{\theta}$ falls
%\end{itemize}
%\end{itemize}
%\bigskip
%\pause
%Note: mean of $\bar{\theta}$ is unknown -- hard to derive a globally optimal strategy
%%SAY: exploration -- exploitation even in this simple setting
%\end{frame}

@@

\begin{center}
\vspace{-20pt}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]
  \node at (0,0) (source) {$\thetac_j$} ;
  \node[ssum, right of = source, scale = 0.75] (sum1) {+};
  \node[above of = sum1, node distance = 1cm] (z1) {$Z_{1}$};
  \node[right of  = z1, node distance = 4cm]  {
 $Z_i  \overset{\mathrm{i.i.d}}{\sim}\mathcal N(0,\sigma^2)$ };
  \node[int1, right of = sum1, node distance = 2cm, scale = 0.75] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (sum1); 
\draw[->,line width = 2pt] (sum1) -- node[above] {$X_{1}$} (enc1); 
\draw[->,line width = 2pt] (z1) -- (sum1);

@@

%\node[int1, right of = enc2, node distance = 6cm, scale = 0.75] (est) {Est};
\draw[->,line width = 2pt] (enc1) -- node[above, xshift = 0cm, rotate = -15] (mes1) {one bit} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm, rotate = 12] (mes2) {one bit} (est);   
\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\thetac}_j$};
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\begin{itemize}
\item $X_{1},\ldots,X_{n} \overset{\mathrm{i.i.d}}{\sim}\mathcal N(\theta,\sigma^2)$
\pause
\item Replicate $k$ times: $\thetac_1,\ldots,\thetac_k \overset{\iid}{\sim} \mathcal N(0,\sigma_0^2)$ 
\pause
\item Block encode $X_{j,1},\ldots,X_{j,n}$ using $k$ bits, $j=1,\ldots,k$
\pause
\item $D_{CEO} = \underset{k,\widehat{\theta}}{\inf} \frac{1}{k} \sum_{j=1}^k \mathbb E \left( \theta_j - \widehat{\theta}_j\right)^2$
\end{itemize}
%SAY: lower bound to our setting
\end{frame}

\begin{frame}
Quadratic Gaussian CEO under optimal rate allocation [Chen, Zhang, Berger, Wicker '04] :
\[
 D_{CEO}  \geq \frac{4}{3} \frac{\sigma^2}{n} + o(1/n) 
\]
%\begin{itemize}
%\item[] CEO distortion satisfies [Chen, Zhang, Berger, Wicker '04]:
%\[ D^\star = \frac{4}{3} %\frac{\sigma^2}{n} + o(n^{-1}) \]
%\end{itemize}
\pause
\begin{alertblock}{Conclusion}
Distributed encoding will hurt you (even if you can repeat experiment and encode over blocks)
\end{alertblock}
\bigskip
%\pause
%\begin{itemize}
%\item In fact [K., Rini, Goldsmith '17]:
%\end{itemize} 
%\[
% D_{CEO} \leq \frac{4}{3} \frac{\sigma^2}{n} + %\frac{\sigma_0^2}{3n} + o(1/n) 
%\]
%(ARE of $4/3$ is achievable with coding over blocks)
%SAY: depending on how large our bound from $k=1$, it tells us that distributed setting is not the limitng factor
\end{frame}

\section{Adaptive Encoding} 
\begin{frame}
\frametitle{Three Encoding Scenarios}

@@

\begin{itemize}
\item Distributed: $M_i = f_i(X_i)$  \hfill{ARE $\geq 4/3$}
\visible<2->{
\item {Centralized}: $M^n = f(X_1,\ldots,X_n)$ \hfill{ARE $=1$} }
\visible<3->{
\item Next: {\color{red} adaptive / sequential: $M_i = f_i(X_i,M^{i-1})$ \hfill{ARE$=\pi/2$} }  }
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item[(i)] For $X \sim \mathcal{N}(\theta,\sigma^2)$, $\mathsf{med}(X) = \theta$ 
\pause
\item[(ii)] $\mathsf{med}(X) = \underset{m}{ \mathrm{argmin~}} \mathbb E \left|X-m \right|$
\pause
\item[(iii)] Stochastic gradient descent on $\mathbb E \left| X-\theta \right|$:
\[
\theta_n = \theta_{n-1} + \gamma_n \sgn(X_n - \theta_n)
\]
\[
\widehat{\theta}_n = \frac{1}{n} \sum_{i=1}^n \theta_i 
\]
\end{itemize}
\bigskip
\pause
From [Polyak \& Juditsky '92] (under conditions on $(\gamma_n)$): 
\[
\sqrt{n}(\theta- \widehat{\theta}_n) \rightarrow \mathcal N\left(0, \sigma^2 \pi /2 \right)
\]
(MSE convergence follows from [Polyak '90]) 
\end{frame}
%picture

@@

\begin{frame}
\frametitle{Converse}
\framesubtitle{ARE $\geq \pi/2$}
Let $\pi(\theta)$ be a prior on $\Theta$. The van-Trees inequality (e.g. [Tsybakov '08]) implies
\[
\mathbb E \left(\theta - \widehat{\theta} \right)^2 \geq \frac{1}{\mathbb E I_\theta(M^n) + I_\pi} \geq \frac{1}{\sum_{i=1}^n I_\theta(M_i|M^{i-1}) + I_\pi} 
\]
$I_\pi$ is the location Fisher information with w.r.t. $\pi(\theta)$ \\

@@

\begin{frame}
\frametitle{Distributed Encoding}
\framesubtitle{Threshold Detection}
We consider only messages of the form
\[
M_i = \sgn(X_i - t_i),\quad i=1,\ldots,n
\]
\begin{center}
\pause
\begin{tikzpicture}
\node[coordinate] at(0,0) (ti) {};
\draw[color = red, line width = 1.5pt] (ti)+(0,0.2) -- node[above, yshift = 0.08cm] {\color{black} $t_i$} +(0,-0.2);
\draw[<-,line width = 1pt] (-3,0) -- node[above] {$M_i=-1$} (ti);
\draw[->,line width = 2pt] (ti) -- node[above] {$M_i=1$} (6,0);
\end{tikzpicture}
\end{center}
Assume:
\[
\lambda_n([a,b]) = \frac{1}{n} \mathsf{card}\left( [a,b]  \cap \left\{t_i \right\} \right)
\]
converges weakly to a probability distribution $\lambda$
\\
\bigskip
\pause
\textbf{Example:} $t_1,\ldots,t_n$ drawn i.i.d. from a distribution $\lambda$ on $\mathbb R$

@@

\begin{frame}
\frametitle{Interpretations}
\begin{itemize}
\pause
\item MLE is local asymptotically minimax 
\pause
\item ARE of MLE is $1/K_{\lambda}(\theta)$ -- 
only depends on asymptotic threshold density $\lambda$
\pause
\item  
\[
\mathrm{ARE} = \frac{1}{K_{\lambda}(\theta)}  = \frac{\sigma^2}{\int \eta \left( \frac{t-\theta}{\sigma}\right) \lambda(dt)}
> \pi / 2
\]
(equality iff $\lambda = \delta_{\theta}$)
%SAY: cannot be as good as sequential thresholds
%\item Threshold density $\lambda$ that maximizes
\pause
\item Minimax $\lambda$ can be found using a convex program -- minimax ARE  depends on radius of $\Theta$
%ARE depends on size of parameter space $\Theta$ 
%\begin{center}
%\begin{tikzpicture}
%\node at (0,0)  {\includegraphics[scale=0.3]{k_vs_b2}};
%\node at (0,-1.7) {$r$};
%\node[rotate = 90] at (-2.3,0) {$1/K_{\lambda^\star}$};
%\end{tikzpicture}
%
%Minimax ARE vs radius $r$ of parameter space
%\end{center}

@@

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\pause
\item Centralized encoding: ARE$=1$
\pause
\item Adaptive:
\begin{itemize}
\item ARE is $\pi/2$ (regardless of radius of $\Theta$)
\pause
\item $\sim 1.57$ more samples are required due to 1-bit constraints
\end{itemize}
\pause
%\pause
%\item One-step optimal one-bit message is a threshold detector
\bigskip
\item Distributed threshold detection:
\begin{itemize}
\item ARE of MLE characterized by density of threshold values 
\pause
\item MLE is local asymptotically optimal 
\pause
\item minimax ARE depends on radius of $\Theta$
\end{itemize}
%\item Minimax ARE of MLE depends on size of parameter space
\end{itemize}

@@

\begin{frame}
\frametitle{One-step Optimal Scheme}
Initialization: $P_0(t) = \pi(\theta)$ \\
Repeat for $n \geq 1$:
\begin{itemize}
\item[(i)] $P_n(t) = \Prob(\theta = t|M^n) = \alpha_n P_{n-1}(t) \Phi \left(M_n \frac{t-\tau_{n-1}}{\sigma} \right)$ 
\item[(ii)] $\widehat{\theta} = \mathbb E[\theta | M^n] = \int t P_n(t)dt $ 
\item[(iii)] Find $\tau_n$ from 
\[
\tau_n = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau t P_n(t)dt }{\int_{-\infty}^\tau P_n(t)dt}  + \frac{ \int_{\tau}^\infty t P_n(t)dt}{\int_{\tau}^\infty P_n(t)dt} \right)
\]
\item[(iv)] $M_{n+1} = \sgn(X_{n+1}-\tau_n)$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Numerical Example}
\begin{center}
Normalized empirical risk versus number of samples $n$ \\
(1000 Monte Carlo experiments)
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.5]{one_bit_adaptive}};
%\node[color = blue,  minimum width=1cm, align = left] at (1.5,1.5) {assymptotically optimal \eqref{eq:sgd_est}};
%\node[color = red, minimum width=1cm, align = left] at (1.5,0.55) {one-step optimal \eqref{eq:message_update}} ;
\node[rotate = 90, scale = 1] at (-5,0) {$n \mathbb E \left(\widehat{\theta}_n - \theta \right)^2$};
\node[scale = 1] at (0,-3) {$n$};
%\draw[dashed] (-3.5,1.18) node[left, scale = 0.7] {$\frac{\pi}{2}$}-- (-3,1.18);
\end{tikzpicture}
$\theta \sim \unif(-3,3)$
\end{center}
\end{frame}

### 2017-10-05 Writeups/proof of Lemma 7 for log-concave distributions.tex
new version of presentation

\begin{lem} \label{lem:bound_intervals}
Assume that $X_1,\ldots,X_n$ are distributed with CDF $F(x)$ and PDF $f(x)$ such that $F(-x) = 1-F(x)$ (and $f(-x)=f(x)$). Assume moreover that 
$ f(x) = e^{g(x}}$ where $g(x)$ is concave and twice differentiable such that $g''(x)<0$. Then for any $x_1 \geq \ldots \geq x_n \in \mathbb R$, we have 
\begin{equation}
\frac{ \left(  \sum_{k=1}^n (-1)^{k+1} f(x_k) \right)^2} 
{\left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)  } \leq \frac{2} {\pi}. \label{eq:bound_intervals}
\end{equation}
\end{lem}

\subsubsection*{Proof of Lem.~\ref{lem:bound_intervals}}We use induction on $n \in \mathbb N$. For the base case $n=1$ we have 
\begin{equation} \label{eq:induction_base}
\frac{  f^2(x)} 
{F(x) \left(1 - F(x) \right) }.
\end{equation}
Since $f(x)$ is log-concave, so does $F(x)$. We now write $F(x) = e^{h(x)}$ with $h(x)$ concave. Note that $f(x) = h'(x) e^{h(x)}$, $h'(x)\geq 0$, $f(x) = f(-x) = h'(-x)e^{h(-x)} $, and $h''(x)\leq 0$ ($h(x)$ is twice differentiable since $F(x)$ is twice differentiable). 
Taking the logarithm of \eqref{eq:induction_base} leads to 
\begin{align*}
& 2 \log f(x) - \log F(x) + \log F(-x) \\
& = 2 \log h'(x) + h(x) + h(-x).
\end{align*}
Since $h'(x)$ is non-increasing, the maximal value of \eqref{eq:induction_base} is obtained as 
{HERE!! \color{red} finish proof by showing that \eqref{eq:induction_base} is smaller than $4f(0)$.\\
}
\\
Assume now that \eqref{eq:bound_intervals} holds for all integers up to some $n = N-1$ and consider the case $n = N$. The maximal value of \eqref{eq:bound_intervals} is attained for the same $(x_1,\ldots,x_N) \in \mathbb R^N$ that attains the maximal value of 
\begin{align*}
& g(x_1,\ldots, x_N) \triangleq 2 \log \left(  \sum_{k=1}^{N} (-1)^{k+1} f(x_k) \right) - \\
&  \log
\left( \sum_{k=1}^N (-1)^{k+1} F(x_k) \right)
-\log \left(1 -  \sum_{k=1}^N (-1)^{k+1} F(x_k) \right) \\
& = 2 \log \delta_N - \log \Delta_N - \log \left(1 - \Delta_N  \right),
\end{align*}
where we denoted $\delta_N \triangleq \sum_{k=1}^{N} (-1)^{k+1} f(x_k)$ and $\Delta_N =  \sum_{k=1}^N (-1)^{k+1} F(x_k)$. The derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by
\[
\frac{\partial  g}{\partial x_k} = \frac{2 (-1)^{k+1} f'(x_k)}{\delta_N} -\frac{(-1)^{k+1} f(x_k)}{\Delta_N } + \frac{(-1)^{k+1} \phi(x_k)}{1-\Delta_N }.
\]
We conclude that the gradient of $g$ vanishes only if
\[
\frac{f'(x_k)}{f(x_k)} = g'(x_k) = \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.
\]
Due to our assumption that $g''(x) < 0$, the function $g'(x)$ is strictly decreasing. As a result, $g'(x)$ is one to one and the condition above implies $x_1 = \ldots = x_N$. If $N$ is odd then for $x_1=\ldots =x_N$ we have that the LHS of \eqref{eq:bound_intervals} equals
\[
\frac{f(x_1)^2}{ F(x_1) (1 - F(x_1))},
\]
which was shown to be smaller than $4f^2(0)$. If $N$ is even, then for any constant $c$ the limit of the LHS of \eqref{eq:bound_intervals} as $(x_1,\ldots,x_N)\rightarrow (c,\ldots,c)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:bound_intervals} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:bound_intervals} is maximized at the borders, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches plus or minus infinity. For simplicity we only consider the cases where $x_N$ goes to minus infinity or $x_1$ goes to plus infinity (the general case where the first $m$ coordinates goes to infinity or the last $m$ to minus infinity is obtained using similar arguments). Assume first $x_N \rightarrow -\infty$. Then the LHS of \eqref{eq:bound_intervals} equals
\begin{align*}
\frac{ \left(  \sum_{k=1}^{N-1} (-1)^{k+1}f(x_k) \right)^2} 
{\left( \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k)  \right) } ,
\end{align*}
which is smaller than $2/\pi$ by the induction hypothesis. Assume now that $x_1 \rightarrow \infty$. Then the LHS of
\eqref{eq:bound_intervals} equals
\begin{align*}
& \frac{ \left(  \sum_{k=2}^{N} (-1)^{k+1}f(x_k) \right)^2} 
{\left( 1 + \sum_{k=2}^{N} (-1)^{k+1} F(x_k) \right)\left(1- 1 - \sum_{k=2}^{N} (-1)^{k+1} F(x_k)  \right) }  \\
& = \frac{ \left(  -\sum_{m=1}^{N} (-1)^{m+1}f(x'_m) \right)^2} 
{\left( 1 - \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m}) \right)\left( \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m})  \right) },
\end{align*}
where $x'_{m} = x_{m+1}$. The last expression is also smaller than $4f(0)$ by the induction hypothesis. This proves Lem.~\ref{lem:bound_intervals}. \\

>>>>

\begin{lem} \label{lem:bound_intervals}
Let $f(x)$ be a log-concave, symmetric, and twice differentiable pdf. Denote by $F(x)$ the corresponding CDF. Then for any $x_1 \geq \ldots \geq x_n \in \mathbb R$, we have 
\begin{equation}
\frac{ \left(  \sum_{k=1}^n (-1)^{k+1} f(x_k) \right)^2} 
{\left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)  } \leq \frac{2} {\pi}. \label{eq:bound_intervals}
\end{equation}
\end{lem}

\subsubsection*{Proof of Lem.~\ref{lem:bound_intervals}}We use induction on $n \in \mathbb N$. For the base case $n=1$ we have 
\begin{equation} \label{eq:induction_base}
\eta(x) \triangleq \frac{  f^2(x)} 
{F(x) \left(1 - F(x) \right) }.
\end{equation}
Since $f(x)$ is log-concave, positive, and twice differentiable, we may write $f(x) = e^{g(x)}$ with $g(x)$ concave and twice differentiable. Moreover, since symmetry of $f(x)$ implies symmetry of $g(x)$, and, since the latter is concave, it attains a single maximum at the symmetry point $x=0$. Therefore, $x=0$ is the only point for which $g'(x)=0$. We now show that $x=0$ is the point where $\eta(x)$ attains its maximum value. Since $\eta(x)$ is positive and  goes to zero as $x\rightarrow \pm \infty$, the maximum of $\eta(x)$ cannot be attained as $x\rightarrow \pm \infty$. We therefore consider the derivative of $\eta(x)$, given as
\[
\eta'(x) = 2 \eta(x) e^{g(x)} \left(\frac{F(x)-1/2}{F(x)(1-F(x))} + g'(x) \right). 
\]
Since $f(x)$ is symmetric and integrates to $1$, we 
have that $F(0)=0.5$ and hence $\eta'(0) = 0$.  It is left to verify that $\eta'(x)\neq 0$ for $x \neq 0$.

Taking the logarithm of \eqref{eq:induction_base} leads to 
\begin{align*}
& 2 \log f(x) - \log F(x) - \log F(-x) \\
& = 2 \log h'(x) + h(x) - h(-x).
\end{align*}
Note that since $F(x)$ is non-decreasing, so is $h(x)$. In addition, $h'(x)$ is non-increasing since $h(x)$ 
, the maximal value of \eqref{eq:induction_base} is obtained as 
{HERE!! \color{red} finish proof by showing that \eqref{eq:induction_base} is smaller than $4f(0)$.\\
}
\\
Assume now that \eqref{eq:bound_intervals} holds for all integers up to some $n = N-1$ and consider the case $n = N$. The maximal value of \eqref{eq:bound_intervals} is attained for the same $(x_1,\ldots,x_N) \in \mathbb R^N$ that attains the maximal value of 
\begin{align*}
& g(x_1,\ldots, x_N) \triangleq 2 \log \left(  \sum_{k=1}^{N} (-1)^{k+1} f(x_k) \right) - \\
&  \log
\left( \sum_{k=1}^N (-1)^{k+1} F(x_k) \right)
-\log \left(1 -  \sum_{k=1}^N (-1)^{k+1} F(x_k) \right) \\
& = 2 \log \delta_N - \log \Delta_N - \log \left(1 - \Delta_N  \right),
\end{align*}
where we denoted $\delta_N \triangleq \sum_{k=1}^{N} (-1)^{k+1} f(x_k)$ and $\Delta_N =  \sum_{k=1}^N (-1)^{k+1} F(x_k)$. The derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by
\[
\frac{\partial  g}{\partial x_k} = \frac{2 (-1)^{k+1} f'(x_k)}{\delta_N} -\frac{(-1)^{k+1} f(x_k)}{\Delta_N } + \frac{(-1)^{k+1} \phi(x_k)}{1-\Delta_N }.
\]
We conclude that the gradient of $g$ vanishes only if
\[
\frac{f'(x_k)}{f(x_k)} = g'(x_k) = \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.
\]
Due to our assumption that $g''(x) < 0$, the function $g'(x)$ is strictly decreasing. As a result, $g'(x)$ is one to one and the condition above implies $x_1 = \ldots = x_N$. If $N$ is odd then for $x_1=\ldots =x_N$ we have that the LHS of \eqref{eq:bound_intervals} equals
\[
\frac{f(x_1)^2}{ F(x_1) (1 - F(x_1))},
\]
which was shown to be smaller than $4f^2(0)$. If $N$ is even, then for any constant $c$ the limit of the LHS of \eqref{eq:bound_intervals} as $(x_1,\ldots,x_N)\rightarrow (c,\ldots,c)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:bound_intervals} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:bound_intervals} is maximized at the borders, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches plus or minus infinity. For simplicity we only consider the cases where $x_N$ goes to minus infinity or $x_1$ goes to plus infinity (the general case where the first $m$ coordinates goes to infinity or the last $m$ to minus infinity is obtained using similar arguments). Assume first $x_N \rightarrow -\infty$. Then the LHS of \eqref{eq:bound_intervals} equals
\begin{align*}
\frac{ \left(  \sum_{k=1}^{N-1} (-1)^{k+1}f(x_k) \right)^2} 
{\left( \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k)  \right) } ,
\end{align*}
which is smaller than $2/\pi$ by the induction hypothesis. Assume now that $x_1 \rightarrow \infty$. Then the LHS of
\eqref{eq:bound_intervals} equals
\begin{align*}
& \frac{ \left(  \sum_{k=2}^{N} (-1)^{k+1}f(x_k) \right)^2} 
{\left( 1 + \sum_{k=2}^{N} (-1)^{k+1} F(x_k) \right)\left(1- 1 - \sum_{k=2}^{N} (-1)^{k+1} F(x_k)  \right) }  \\
& = \frac{ \left(  -\sum_{m=1}^{N} (-1)^{m+1}f(x'_m) \right)^2} 
{\left( 1 - \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m}) \right)\left( \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m})  \right) },
\end{align*}
where $x'_{m} = x_{m+1}$. The last expression is also smaller than $4f(0)$ by the induction hypothesis. This proves Lem.~\ref{lem:bound_intervals}. \\
